A Multi-Modal Model of Object Deformation under Robotic Pushing

### 2016-12-25 twostagedo-rev-Dec-16.tex
Final revisions from Jeremy!

@@

\begin{figure*}[t]
\centering
%\setlength\fboxsep{0pt}    % padding
%\setlength\fboxrule{0.5pt} % line width
%\fbox{
\includegraphics[width=0.8\textwidth]{figures/classification}%}
\caption{Prediction sequences for several predictors--each of which has been learned for a different material--can be compared against real data. The most accurate predictor at any one moment gives the classification for the material type. The notation is explained in the text.}
\label{fig:classification}
\end{figure*}
%In each case the resulting learned models can be run in two modes, prediction and classification. In prediction mode the models make predictions based on a sequence of robot finger positions. To create the prediction for time $t+1$ the previous predicted state at time $t$ is required: thus the predictions are recursive. This enables predictions over arbitrarily long time periods, but makes learning a good predictor challenging. In filtering mode the predictions of different models are compared to the actual outcomes, and this is used to classify the object as being one of the modular models in the set of models. Since the model is multi-modal we can compare both force and deformation data to the corresponding model predictions.

%The training of the model is performed offline using ground truthed data. %The overall algorithm is given in \alref{alg:overall}.
For clarity, the algorithm is described in exactly these three phases: training, simulation (i.e. the prediction phase), and classification. In empirically evaluating the model new data was used, acquired from new finger pushes. We now proceed to describe the shape predictor SP (Section~\ref{sec:model}), which is based on a mass-spring system, and then the force predictor (FP), which is a regression model (Section~\ref{sec:forces}).

\section{A Mass-Spring Model of Deformation}

A mass-spring model is a very well established abstraction, commonly used to simulate the behaviour of certain types of deformable objects.  As a physics based model, once its parameters are acquired, its behaviour is determined by the evolution of a set of differential equations.  Therefore, by offering a method to automatically calibrate those parameters, this model can make long term predictions about the behaviour of deformable objects, under previously unseen interactions, if only the externally applied forces are known.  The range of validity of these predictions greatly depends on the quality of the calibration, but also on the suitability of the mass-spring model as a model of the particular object.  For this reason, there are variations that address specific issues, like \cite{Bourguignon2000} and \cite{Teschner2004}.  In this work, we focus on the machinery required to achieve this automatic calibration of a physics based model and its integration into a system that exploits it to solve prediction and classification tasks.  We emphasize that the physics based model could be easily substituted by another equivalent model.  For example, the same machine will work if another type of mass-spring model is used, or a FEM model, even though numerical results will differ to some extent.  For these experiments we chose to use a slightly modified version of the mass-spring model proposed by Teschner in \cite{Teschner2004}.

@@

The mass-spring model proposed by Teschner et al.\ in \cite{Teschner2004}, which is explained below, was reduced to a 2D model, so that the predictions could be compared with 2D images of the deformable object.  The main idea being that an agent can make rough real time predictions of what it can see and touch and use those predictions to solve various tasks, like classification or grasping.  Here, the surface of the deformable object is approximated by a regular triangular mesh of unitary masses and springs.  \comment{Since the location of the masses and direction of the springs will bias the deformation response of the object, it is important to distribute the masses uniformly and place the springs with a symetric arrangement for an isotropic material.  This placement can be done with an algorithm that takes into account the symmetries of the object at the begining of the experiment.  If the material gets permanently deformed a properly callibrated physics based model, like the one we propose here, will produce an adequate new arrangement from which new interactions can be studied.}

@@

\subsection{Additional Geometric Constraints}
While the mass-spring model incorporates constraints based on mesh geometry, these are soft constraints implemented in the form of an energy function. Even with several of these terms incorporated, such as a term for preservation of angles, the mass-spring model can cause triangles of the mesh to overlap one another and be reversed or flattened.  \comment{All of these situations are innadecuate since this mesh is used to model one face of a 3D object.}  It is thus necessary to employ additional hard geometric constraints to make the mesh internally consistent.  We achieve this through geometric checking subroutines.  There are two geometric constraints that the mesh is enforced to maintain:

@@

\subsection{Tracking to Provide Ground Truth}

\begin{figure}[!t]
\centering
\includegraphics[width=88mm]{arrio2}
\caption{(a) Experimental set up.  A robot finger pushes a block of deformable material sitting on a table.  A camera registers the events of the plane where the main deformations take place.  (b) Evaluation function.  Left top: Mesh over real block of material.  Left bottom: The typical error of the model vs. the ground truth (red) and the typical error of the tracker (blue).  Right top: tracked ground data.  Middle: TP in yellow, FP in blue, FN in red and TN in gray.  Bottom: simulated mesh. }
\label{fig:scene}
\end{figure}

 The contour of the deformable object is used to obtain the ground truth for the training phase, which is the area enclosed by the contour. Since we only require a tracking algorithm which is good enough to evaluate the predictions of the mass-spring system, a very simple one was used.  Geometrically, the contour is represented by a polygon with hundreds of sides, which can also be called a \textit{linear spline} \cite{Menet1991}. The vertices of the polygon work as control points which can be moved to adjust the polygon to the shape of the object as it gets deformed.  Even though this implementation is inspired by the theory of active contours, it works with a much simpler and faster, but less accurate, algorithm, instead of minimizing a global energy.  \comment{This causes some shadowed corners of the material to be missed at some frames, but as \fref{fig:scene}(b) at its bottom left shows, these errors are not significant in comparison to the errors made by the mass-spring model.} This is explained below.  The tracking of the deformable target in the ground truth has two stages:
\begin{enumerate}
 \item The target object is segmented from the environment.  The Canny algorithm is used to extract the edges of the image.  Since the object of interest has a distinctive colour, \comment{its characteristic hue range is extracted from the area enclosed by the initial contour.  Then} the hue of the pixels on every edge are used to keep only edges that could belong to it.  If the average hue of the pixels on an edge falls outside the calibrated range, the edge is eliminated.
 \item The internal representation of its contour, the linear snake, is adapted to its new shape.  The algorithm assumes that no occlusions take place and that the contour does not bend over itself.
\end{enumerate}
For the first frame, the snake is initialised as a rectangle around the target. This rectangle fragments itself into smaller pieces. Each new vertex adheres itself to the closest point on an edge, as reached by growing rings of pixels around its current position, up to a certain threshold distance [\fref{fig:scene}(a)].  If there is no edge within that area, the control point remains where it is. This criterion was selected to make the representation more robust against edges that are easier to detect on some frames than in others.  Once the edge reappears, the control point adapts itself again.  Also, the distance between two consecutive control points is kept within the interval $[min\_distance, max\_distance]$ and two adjacent edges do not form angles smaller than $min\_angle$.  See \alref{alg:regularise}.  This allows the snake to adapt very quickly to the complexity of the contour it represents.

\begin{figure}
 \begin{algorithmicieee}{Regularise Linear Spline}\label{alg:regularise}
 \STATE Let $L$ be a linear spline representing the contour of a $2D$ shape, and $L_i$ the $i^{th}$ control point in a cyclic order.
 \FORALL {$L_i$ in $L$}
 \IF {$dist(L_i,L_{i-1}) > max\_distance$}
  \STATE Insert $midpoint(L_i,L_{i-1})$ after $L_i$
 \ELSIF {$dist(L_i,L_{i-1}) < min\_distance$}
  \STATE erase $L_i$
 \ELSIF {$angle(L_{i-1},L_i,L_{i+1}) < min\_angle$}
  \STATE erase $L_i$
 \ENDIF
 \ENDFOR
\end{algorithmicieee}
\caption{Algorithm used to regularise the number of elements of a linear spline, in accordance with the level of detail required.}
\end{figure}

@@

\section{Learning to predict forces}
%We used two methods to predict the forces.  The first one consists in measuring the total force acting on the vertex of the mesh closest to the actuator, according to the mass-spring model, after solving detected collisions.  This force is feed as input for the following simulation step.
\label{sec:forces}

Additionally and independently from the mass-spring system,  it is possible to learn to make predictions about the expected reaction force of the material, by using information from the stress-strain diagram obtained from the same training data set as the mass-spring model.  This diagram is obtained from the video recorded by the visual system of the robot and the forces recorded by a sensor located at the wrist.  When the readings of the force sensor start to increase and the figer collides with the block of material, the point of zero deformation is identified.  From here on, the displacement of the finger is considered equivalent to the amount of deformation in the material.  Even though this method is not as rigorous as an engineering test, the estimation is good enough for robot manipulation of common materials.

For automatic calibration of the force prediction model for the pushing (as opposed to retraction) phase both families of lines (logarithmic curves and straight lines) are considered.  To determine the points of discontinuity between segments, each family is evaluated incrementally. The candidate line for a particular family is obtained by fitting a growing subset of data points and calculating the fitting error.  We begin with a small number of points (e.g. five). To detect the discontinuity new points are added until the corresponding normalized error increases beyond a threshold value.  This marks the point of discontiuity. Having discovered one best parameterized line for each family, the two resulting candidates are compared. We select the type of curve that covers more points in a single segment.  This model is used to predict the forces for the mass-spring simulation \fref{fig:diagram} and \fref{fig:sstrain}.%The prediction can be averaged with the spring reaction force calculated from the previous step, to combine the visual and haptic modalities and test whether the quality of the prediction is improved.

\section{Experimental Results}
\label{sec:experiments}
\subsection{Scenario}
For the experimental scenario we used two materials: sponge and plasticine. Each formed a rectangular block of deformable material, of $8.5 \unit{cm}$ long, $5 \unit{cm}$ wide and $1.8 \unit{cm}$ depth. One of the objects is selected, and is placed on a table; the robot pushes the object with a cylindrical finger, followed by a retrieval movement.  A colour firewire camera records the action perpendicularly, while a force sensor registers the reaction forces in synchrony with each photograph taken [\fref{fig:scene}(a)].  \comment{This set up allowed us to study the effect of using an automatically calibrated physics based model as model for the material, while using 2D meshes, even though the same principles can be used if we use a 3D point cloud and a 3D mesh.  To generate the initial mesh for the mass-spring model a routine placed the nodes for the masses evenly spaced on a rectangular grid, given the number of squares we wanted per side.  We made experiments with grids of $3x2, 10x5$ and $20x10$ squares.  To aliviate the bias produced by the springs they were placed in a star pattern, thus compensating forces on eight directions.}

@@

\comment{We also compared the generalisation capabilities of our model against the system proposed in \cite{Cretu2012}.  This system uses a Growing Neural Gas (GNG) Network to calibrate a colour segmentation algorithm.  We couldn't use the HSV space as suggested because our videos have lots of shadows which produce holes inside the material regions which cause problems in further stages.  We got better results with Luv color space.  We clarify that, even though we used the HSV space in our tracking algorithm, we didn't have the same problems because the nodes of our linear snake search for nearby edges and are not affected by shadows in the middle of the object or temporary failures in detecting an edge.  Another GNG network is used to select the control points adequate for approximating the contour in the first frame.  Later, Cretu et al. use a Neural Gas (NG) algorithm to adjust that fixed number of control points to the contour of the material at every frame.  Using the coordinates of the finger, the sensed force and the coordinates of those control points through time, they train a Feedforward Neural Network (FFNN) to predict the contour of the material given the position of the finger and sensed force.  We reproduced those steps and their training method on our training video.  However, when we traid to test the generalisation capabilities of their method to new pushing actions, the weakness of the FFNN method became evident: it got tied to the specific location and type of interaction it was trained with, while our physics based model had no problem at making predictions for new pushing actions.  The results with this other method can be seen on \fref{fig:cretu_prediction}.

Since the videos with the plasticine have even greater shadows it was not posible to use their tracking algorithm on them.  However, even if we could change that algorithm, we would expect the system to have the same problems and predict the deformation at the middle of the material even if is being pushed somewhere else. }

@@

\begin{table}[!t]
\renewcommand{\arraystretch}{1.3}
\caption{Classification per Frame}
\label{tab:classperframe}
\centering
\begin{tabular}{ccccccccc}
\hline
 & & \multicolumn{3}{c}{\bfseries Sponge} & & \multicolumn{3}{c}{\bfseries Plasticine} \\
Data Set & & TP & FN & Precision & & TP & FN & Precision\\
\hline\hline
 \multicolumn{9}{c}{Trained with forces on vertices and no collision resolution.} \\
\hline
   training & & 167 & 132 & 0.56 & & 186 & 113 & 0.62 \\
   test1 & & 167 & 132 & 0.56 & & 271 & 28 & 0.91 \\
   test2 & & 242 & 57 & 0.81 & & 289 & 10 & 0.97 \\ 
\hline
\multicolumn{9}{p{0.9\columnwidth}}{\begin{footnotesize}\comment{* TP and FN measured in number of frames. TP + FN = total number of frames.  Final class assigned according to the performance of the models' predictions across the whole video.} \end{footnotesize}}
\end{tabular}
\end{table}
% \begin{table}[!t]
% \renewcommand{\arraystretch}{1.3}
% \caption{Classification per Frame}
% \label{tab:classperframe}
% \centering
% \begin{tabular}{ccccccccc}
% \hline
%  & & \multicolumn{4}{c}{\bfseries Sponge} \\
% Data Set & & TP & FN & TN & FP & Precision \\
% \hline\hline
%  \multicolumn{9}{c}{Trained with forces on vertices and no collision resolution.} \\
% \hline
%    training & & 167 & 132 & & & 0.56  \\
%    test1 & & 167 & 132 & & & 0.56  \\
%    test2 & & 242 & 57 & & & 0.81 \\ 
% \hline
% %\multicolumn{7}{p{68mm}}{\begin{footnotesize}* TP and FN measured in number of frames.  Final class assigned according to the performance of the models' predictions across the whole video. \end{footnotesize}}
% \end{tabular}
% 
% \vspace*{1em}
% \begin{tabular}{ccccccccc}
% \hline
%  & & \multicolumn{4}{c}{\bfseries Plasticine} \\
% Data Set & & TP & FN & TN & FP & Precision\\
% \hline\hline
%  \multicolumn{9}{c}{Trained with forces on vertices and no collision resolution.} \\
% \hline
%    training & & 186 & 113 & & & 0.62 \\
%    test1 & & 271 & 28 & & & 0.91 \\
%    test2 & & 289 & 10 & & & 0.97 \\ 
% \hline
% %\multicolumn{7}{p{68mm}}{\begin{footnotesize}* TP and FN measured in number of frames.  Final class assigned according to the performance of the models' predictions across the whole video. \end{footnotesize}}
% \end{tabular}
% \end{table}

\comment{
\tref{tab:classperframe} shows the results for the classification task, frame per frame, for the videos that were used as training and test sets for the prediction task.  When the sponge model performed better for a frame of the sponge video it is marked as a true positive, when plasticine does better it is marked as a false negative (since it didn't recognize it as a sponge).  The same process is used for the other material.  Since the concept was proved with only two materials, true positives and false negatives provide all the information.

It calls for attention that in some of the experiments, like the one illustrated above, the performance of classification for the sponge on the training and test1 sets are the same.  We interpret that the mass-spring physics based model generalizes naturally when the force/deformation conditions are similar, even if the place for pushing is different.  Between pushing in the middle and pushing closer to the border the qualitative behaviour of the elastic material was no so different for these models; unlike the data from the third set, where the distribution of mass and maximum force/deformation is more different from the ones seen on the training set of the model.  It is interesting to confirm that the difference in predictions between the models is bigger here in favour of a better classification.

It is also remarkable that, even though, when modeling the sponge, the parameters for the plasticine were close in performance to the parameters for sponge (thus causing more problems to establish a good classification), the opposite does not happen.  That is, the parameters for the sponge are not good at modeling the plasticine on the test sets, thus yielding better results at the classification task.  This could happen because the plastic behaviour (that is, the permanent deformation of a spring) gets activated with stronger forces after the $yield$ limit is reached, which only appear when pushing the plasticine.  With the small forces feedback by the sponge, the model for the plasticine does not reach its range of plastic behaviour and thus the big differences in behaviour do not manifest.
}

@@

There is considerable future work to be done. The first extension is due to the fact that we restricted the deformation model to 2D, and this can easily revert to the 3D meshes as used in Teschner's original work. The second issue is to explore new methods for the learning phase. The current search procedure doesn't always find good solutions for the plastic deformation model. But a more basic problem is that mass-spring models restrict the range of deformations that can be modelled. While the framework presented here is promising, we regard the actual deformation model representation as an open research problem. \comment{Finally, all the work here is for off-line learning, from a carefully chosen viewpoint. The issue of online learning of these representations, including how to explore so as to learn, is an open issue.}

>>>>

\begin{figure*}[!t]
\centering
\includegraphics[width=0.8\textwidth]{figures/classification}
\caption{Prediction sequences for several predictors--each of which has been learned for a different material--can be compared against real data. The most accurate predictor at any one moment gives the classification for the material type. The notation is explained in the text.}
\label{fig:classification}
\end{figure*}

@@

%In each case the resulting learned models can be run in two modes, prediction and classification. In prediction mode the models make predictions based on a sequence of robot finger positions. To create the prediction for time $t+1$ the previous predicted state at time $t$ is required: thus the predictions are recursive. This enables predictions over arbitrarily long time periods, but makes learning a good predictor challenging. In filtering mode the predictions of different models are compared to the actual outcomes, and this is used to classify the object as being one of the modular models in the set of models. Since the model is multi-modal we can compare both force and deformation data to the corresponding model predictions.

%The training of the model is performed offline using ground truthed data. %The overall algorithm is given in \alref{alg:overall}.
For clarity, the algorithm is described in exactly these phases: training, simulation (i.e. the prediction phase), and classification. In empirically evaluating the model new data was used, acquired from new finger pushes. We now proceed to describe first the visual analysis and force measurement, the force predictor (FP), which is a regression model (Section~\ref{sec:forces}), and then the shape predictor SP (Section~\ref{sec:model}), which is based on a mass-spring system.

\section{Visual Analysis}
\label{sec:vision}

Although the visual analysis is not an original contribution of the paper, it is necessary to track the shape of the object as it deforms. This tracking is performed on the two training movies to provide training data for the shape predictor (SP). For the test movies the system only requires visual analysis of the first frame to initialise the predictors. The test movies are analysed for each frame, but only for experimental evaluation purposes. We now give details of the visual analysis method.

\subsection{Tracking to Provide Ground Truth}
\begin{figure}[!t]
 \begin{algorithmicieee}{Regularised Linear Spline}\label{alg:regularise}
 \STATE Let $L$ be a linear spline representing the contour of a $2D$ shape, and $L_i$ the $i^{th}$ control point in a cyclic order.
 \FORALL {$L_i$ in $L$}
 \IF {$dist(L_i,L_{i-1}) > max\_distance$}
  \STATE Insert $midpoint(L_i,L_{i-1})$ after $L_i$
 \ELSIF {$dist(L_i,L_{i-1}) < min\_distance$}
  \STATE erase $L_i$
 \ELSIF {$angle(L_{i-1},L_i,L_{i+1}) < min\_angle$}
  \STATE erase $L_i$
 \ENDIF
 \ENDFOR
\end{algorithmicieee}
\caption{Algorithm used to regularise the number of elements of a linear spline, in accordance with the level of detail required.}
\end{figure}
 The contour of the deformable object is used to obtain the ground truth for the training phase, which is the area enclosed by the contour. Since we only require a tracking algorithm which is good enough to evaluate the predictions of the mass-spring system, a very simple one was used.  Geometrically, the contour is represented by a polygon with hundreds of sides, which can also be called a \textit{linear spline} \cite{Menet1991}. The vertices of the polygon work as control points which can be moved to adjust the polygon to the shape of the object as it gets deformed.  Even though this implementation is inspired by the theory of active contours, it works with a much simpler and faster, but less accurate, algorithm, instead of minimizing a global energy.  \comment{This causes some shadowed corners of the material to be missed at some frames, but as \fref{fig:scene}(b) at its bottom left shows, these errors are not significant in comparison to the errors made by the mass-spring model.} This is explained below.  The tracking of the deformable target in the ground truth has two stages:
 \begin{figure}[t]
\centering
\includegraphics[width=88mm]{arrio2}
\caption{(a) Experimental set up.  A robot finger pushes a block of deformable material sitting on a table.  A camera registers the events of the plane where the main deformations take place.  (b) Evaluation function.  Left top: Mesh over real block of material.  Left bottom: The typical error of the model vs. the ground truth (red) and the typical error of the tracker (blue).  Right top: tracked ground data.  Middle: TP in yellow, FP in blue, FN in red and TN in gray.  Bottom: simulated mesh. }
\label{fig:scene}
\end{figure}

\begin{enumerate}
 \item The target object is segmented from the environment.  The Canny algorithm is used to extract the edges of the image.  Since the object of interest has a distinctive colour, \comment{its characteristic hue range is extracted from the area enclosed by the initial contour.  Then} the hue of the pixels on every edge are used to keep only edges that could belong to it.  If the average hue of the pixels on an edge falls outside the calibrated range, the edge is eliminated.
 \item The internal representation of its contour, the linear snake, is adapted to its new shape.  The algorithm assumes that no occlusions take place and that the contour does not bend over itself.
\end{enumerate}
For the first frame, the snake is initialised as a rectangle around the target. This rectangle fragments itself into smaller pieces. Each new vertex adheres itself to the closest point on an edge, as reached by growing rings of pixels around its current position, up to a certain threshold distance [\fref{fig:scene}(a)].  If there is no edge within that area, the control point remains where it is. This criterion was selected to make the representation more robust against edges that are easier to detect on some frames than in others.  Once the edge reappears, the control point adapts itself again.  Also, the distance between two consecutive control points is kept within the interval $[min\_distance, max\_distance]$ and two adjacent edges do not form angles smaller than $min\_angle$.  See \alref{alg:regularise}.  This allows the snake to adapt very quickly to the complexity of the contour it represents.

\section{Learning to predict forces}
\label{sec:forces}

It is possible to learn the force predictor (FP), by using information from the stress-strain diagram obtained from the data set of measured positions and forces.  This diagram is obtained from the video recorded by the visual system of the robot and the forces recorded by the sensor located at the wrist.  When the readings of the force sensor start to increase and the figer collides with the block of material, the point of zero deformation is identified.  From here on, the displacement of the finger is considered equivalent to the amount of deformation in the material.  Even though this method is not as rigorous as an engineering test, the estimation is good enough for robot manipulation of common materials.

For automatic calibration of the force prediction model for the pushing (as opposed to retraction) phase both families of lines (logarithmic curves and straight lines) are considered.  To determine the points of discontinuity between segments, each family is evaluated incrementally. The candidate line for a particular family is obtained by fitting a growing subset of data points and calculating the fitting error.  We begin with a small number of points (e.g. five). To detect the discontinuity new points are added until the corresponding normalized error increases beyond a threshold value.  This marks the point of discontiuity. Having discovered one best parameterized line for each family, the two resulting candidates are compared. We select the type of curve that covers more points in a single segment.  This model is used to predict the forces for the mass-spring simulation \fref{fig:diagram} and \fref{fig:sstrain}.

\section{A Mass-Spring Model of Deformation}
The shape predictor (SP) is based on a mass-spring model of deformation.  A mass-spring model is a very well established abstraction, commonly used to simulate the behaviour of certain types of deformable objects.  As a physics based model, once its parameters are acquired, its behaviour is determined by the evolution of a set of differential equations.  Therefore, by offering a method to automatically calibrate those parameters, this model can make long term predictions about the behaviour of deformable objects, under previously unseen interactions, if only the externally applied forces are known.  The range of validity of these predictions greatly depends on the quality of the calibration, but also on the suitability of the mass-spring model as a model of the particular object.  For this reason, there are variations that address specific issues, like \cite{Bourguignon2000} and \cite{Teschner2004}.  In this work, we focus on the machinery required to achieve this automatic calibration of a physics based model and its integration into a system that exploits it to solve prediction and classification tasks.  We emphasize that the physics based model could be easily substituted by another equivalent model.  For example, the same machine will work if another type of mass-spring model is used, or a FEM model, even though numerical results will differ to some extent.  For these experiments we chose to use a slightly modified version of the mass-spring model proposed by Teschner in \cite{Teschner2004}.

@@

The mass-spring model proposed by Teschner et al.\ in \cite{Teschner2004}, which is explained below, was reduced to a 2D model, so that the predictions could be compared with 2D images of the deformable object.  The main idea being that an agent can make rough real time predictions of what it can see and touch and use those predictions to solve various tasks, like classification or grasping.  Here, the surface of the deformable object is approximated by a regular triangular mesh of unitary masses and springs.  \comment{Since the location of the masses and direction of the springs will bias the deformation response of the object, it is important to distribute the masses uniformly and place the springs with a symetric arrangement for an isotropic material.  This placement can be done with an algorithm that takes into account the symmetries of the object at the begining of the experiment.  If the material gets permanently deformed a properly calibrated physics based model, like the one we propose here, will produce an adequate new arrangement from which new interactions can be studied.}

@@

\subsection{Additional Geometric Constraints}
While the mass-spring model incorporates constraints based on mesh geometry, these are soft constraints implemented in the form of an energy function. Even with several of these terms incorporated, such as a term for preservation of angles, the mass-spring model can cause triangles of the mesh to overlap one another and be reversed or flattened.  \comment{All of these situations are inadequate since this mesh is used to model one face of a 3D object.}  It is thus necessary to employ additional hard geometric constraints to make the mesh internally consistent.  We achieve this through geometric checking subroutines.  There are two geometric constraints that the mesh is enforced to maintain:

@@

@@

\section{Experimental Results}
\label{sec:experiments}
\subsection{Scenario}
For the experimental scenario we used two materials: sponge and plasticine. Each formed a rectangular block of deformable material, of $8.5 \unit{cm}$ long, $5 \unit{cm}$ wide and $1.8 \unit{cm}$ depth. One of the objects is selected, and is placed on a table; the robot pushes the object with a cylindrical finger, followed by a retrieval movement.  A colour firewire camera records the action perpendicularly, while a force sensor registers the reaction forces in synchrony with each photograph taken [\fref{fig:scene}(a)].  \comment{This set up allowed us to study the effect of using an automatically calibrated physics based model as model for the material, while using 2D meshes, even though the same principles can be used if we use a 3D point cloud and a 3D mesh.  To generate the initial mesh for the mass-spring model a routine placed the nodes for the masses evenly spaced on a rectangular grid, given the number of squares we wanted per side.  We made experiments with grids of $3 \times 2, 10 \times 5$ and $20 \times 10$ squares.  To alleviate the bias produced by the springs they were placed in a star pattern, thus compensating forces on eight directions.}

@@

\comment{We also compared the generalisation capabilities of our model against the system proposed in \cite{Cretu2012}.  This system uses a Growing Neural Gas (GNG) Network to calibrate a colour segmentation algorithm.  We couldn't use the HSV space, as suggested by Cretu, because our videos have lots of shadows which produce holes inside the material regions which cause problems in further stages.  We obtained better results for GNGN with the Luv color space.  Note that, even though we used the HSV space in our tracking algorithm, this didn't have the same problems because the nodes of our linear snake search for nearby edges and are not affected by shadows in the middle of the object or temporary failures in detecting an edge.  Another GNG network is used to select the control points adequate for approximating the contour in the first frame.  Later, Cretu et al. used a Neural Gas (NG) algorithm to adjust that fixed number of control points to the contour of the material at every frame.  Using the coordinates of the finger, the sensed force and the coordinates of those control points through time, they train a Feedforward Neural Network (FFNN) to predict the contour of the material given the position of the finger and sensed force.  We reproduced those steps and their training method on our training video.  However, when we tried to test the generalisation capabilities of their method to new pushing actions, the weakness of the FFNN method became evident: it became tied to the specific location and type of interaction it was trained with, while our physics based model had no problem making predictions for new pushing actions.  The results with the GNGN method can be seen in \fref{fig:cretu_prediction}.

%Since the videos with the plasticine have even greater shadows it was not posible to use their tracking algorithm on them.  However, even if we could change that algorithm, we would expect the system to have the same problems and predict the deformation at the middle of the material even if is being pushed somewhere else. 
}

@@

\begin{table}[!t]
\renewcommand{\arraystretch}{1.3}
\caption{Classification per Frame}
\label{tab:classperframe}
\centering
\begin{tabular}{ccccccccc}
\hline
 & & \multicolumn{3}{c}{\bfseries Sponge} & & \multicolumn{3}{c}{\bfseries Plasticine} \\
Data Set & & TP & FN & Precision & & TP & FN & Precision\\
\hline\hline
 \multicolumn{9}{c}{Trained with forces on vertices and no collision resolution.} \\
\hline
   training & & 167 & 132 & 0.56 & & 186 & 113 & 0.62 \\
   test1 & & 167 & 132 & 0.56 & & 271 & 28 & 0.91 \\
   test2 & & 242 & 57 & 0.81 & & 289 & 10 & 0.97 \\ 
\hline
\multicolumn{9}{p{0.9\columnwidth}}{\begin{footnotesize}\comment{* TP and FN are measured in terms of numbers of frames. TP + FN = 299.  The final class assigned according to the performance of the models' predictions across the whole video. See text for details.} \end{footnotesize}}
\end{tabular}
\end{table}

% \begin{table}[!t]
% \renewcommand{\arraystretch}{1.3}
% \caption{Classification per Frame}
% \label{tab:classperframe}
% \centering
% \begin{tabular}{ccccccccc}
% \hline
%  & & \multicolumn{4}{c}{\bfseries Sponge} \\
% Data Set & & TP & FN & TN & FP & Precision \\
% \hline\hline
%  \multicolumn{9}{c}{Trained with forces on vertices and no collision resolution.} \\
% \hline
%    training & & 167 & 132 & & & 0.56  \\
%    test1 & & 167 & 132 & & & 0.56  \\
%    test2 & & 242 & 57 & & & 0.81 \\ 
% \hline
% %\multicolumn{7}{p{68mm}}{\begin{footnotesize}* TP and FN measured in number of frames.  Final class assigned according to the performance of the models' predictions across the whole video. \end{footnotesize}}
% \end{tabular}
% 
% \vspace*{1em}
% \begin{tabular}{ccccccccc}
% \hline
%  & & \multicolumn{4}{c}{\bfseries Plasticine} \\
% Data Set & & TP & FN & TN & FP & Precision\\
% \hline\hline
%  \multicolumn{9}{c}{Trained with forces on vertices and no collision resolution.} \\
% \hline
%    training & & 186 & 113 & & & 0.62 \\
%    test1 & & 271 & 28 & & & 0.91 \\
%    test2 & & 289 & 10 & & & 0.97 \\ 
% \hline
% %\multicolumn{7}{p{68mm}}{\begin{footnotesize}* TP and FN measured in number of frames.  Final class assigned according to the performance of the models' predictions across the whole video. \end{footnotesize}}
% \end{tabular}
% \end{table}

\comment{
\tref{tab:classperframe} shows the results for the classification task, frame by frame for each video.  When the sponge model performed better for a frame of the sponge video it is marked as a true positive (TP), when plasticine does better it is marked as a false negative (FN) ---since it didn't recognize it as a sponge.  The analogous process is used for the plasticine. The total TP+FN=299 which is the total frames per movie. The overall classification for a movie can be determined by whether TP > FN or vice versa. Using this rule gives a perfect classification of the six movies.

It should be noted that, in some of the experiments, the performance of classification for the sponge on the training and test1 sets are the same.  We interpret that the mass-spring physics based model generalizes naturally when the force/deformation conditions are similar, even if the place for pushing is different.  Between pushing in the middle and pushing closer to the border the qualitative behaviour of the elastic material was no so different for these models; unlike the data from the third set, where the distribution of mass and maximum force/deformation is more different from the ones seen on the training set of the model.  It is interesting to confirm that the difference in predictions between the models is bigger here, resulting in better classification.

It is also notable that, when modeling the sponge, the parameters for the plasticine were close in performance to the parameters for sponge (thus causing more problems to establish a good classification), the opposite does not happen.  That is, the parameters for the sponge are not good at modeling the plasticine on the test sets, thus yielding better results at the classification task.  This could happen because the plastic behaviour (that is, the permanent deformation of a spring) gets activated with stronger forces after the $yield$ limit is reached, which only appear when pushing the plasticine.  With the small forces fed back by the sponge, the model for the plasticine does not reach its range of plastic behaviour and thus the big differences in behaviour are not manifest.
}

@@

There is considerable future work to be done. The first extension is due to the fact that we restricted the deformation model to 2D, and this can easily revert to the 3D meshes as used in Teschner's original work. The second issue is to explore new methods for the learning phase. The current search procedure doesn't always find good solutions for the plastic deformation model. A third extension is to extend the fairly trivial visual system to initialise mass-spring models for objects of abitrary shape. We propose using a Voronoi triangulation rather than a grid. But a more basic problem is that mass-spring models restrict the range of deformations that can be modelled. While the framework presented here is promising, we regard the actual deformation model representation as an open research problem. \comment{Finally, all the work here is for off-line learning, from a carefully chosen viewpoint. The issue of online learning of these representations, including how to explore so as to learn, is an open issue.}
